---
title: "Adjusting for confounding"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
description: "Mantel-Haenszel and regression methods to adjust for confounding"
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(dplyr)
library(epitools)
library(epiDisplay)
knitr::opts_chunk$set(echo = FALSE)
tutorial_options(
  exercise.completion = TRUE
)

mat <- data.frame(
  diabetes = c(
    rep("present", 4),
    rep("absent", 4)
  ),
  age_group = rep(c(
    "under_45",
    "over_45"
  ), 4),
  sae = rep(c(rep(TRUE, 2), rep(FALSE, 2)), 2),
  count = c(
    34, 19, 31, 5,
    19, 8, 325, 42
  )
)

df <- mat[rep(sequence(nrow(mat)), mat[["count"]]), ]
df$count <- NULL
```

## Controlling for confounding

Confounding is generally a nuisance in statistical analysis. 
It is helpful to know how to adjust the observed effects for 
confounding. 

In this short tutorial, you will practice adjusting for confounding
in two ways - stratification and regression. 

### Example data

The simple example data includes 3 variables: 

* age group (over or under 45 years), 
* diabetes status (present or absent),
* SAE (TRUE or FALSE)

We are interested in whether SAE occurrence is associated with age-group
in this analysis. Based on our knowledge of TB treatment SAEs, and after 
reviewing the literature, we believe that diabetes might be a confounder that
we should adjust for. 

Here are the first 10 rows of the example dataset. Note the variable names, 
and the values the different variables can take: 

```{r example-data}
knitr::kable(head(df, 10), row.names = FALSE)
```
The tables below give you a summary of the distribution of the exposure and 
outcome variable by diabetes status: 

#### Diabetics

```{r diabetic-sae-by-age}
dfd <- df[which(df$diabetes == "present"), ]
knitr::kable(table(dfd$age_group, SAE = dfd$sae))
```

#### Non-diabetics

```{r non-diabetic-sae-by-age}
dfd <- df[which(df$diabetes == "absent"), ]
knitr::kable(table(dfd$age_group, SAE = dfd$sae))
```

### Crude Odds Ratio

Can you calculate the odds ratio for SAE occurrence in older 
compared to younger study participants? **Round your result to 2 decimal 
places**

```{r odds-age-outcome, exercise = TRUE, exercise.lines = 10, exercise.eval = TRUE}
head(df)
```
```{r odds-age-outcome-solution}
result <- epitools::oddsratio.wald(
  x = (df$age_group == "over_45"),
  y = df$sae
)

round(result$measure[2],
  digits = 2
)
```
```{r odds-age-outcome-hint }
table(df$age_group, df$sae)
```
```{r odds-age-outcome-check}
result <- epitools::oddsratio.wald(
  x = (df$age_group == "over_45"),
  y = df$sae
)

or <- round(result$measure[2],
  digits = 2
)

grade_this({
  if (length(.result) > 1) {
    fail("You need to enter a single number, rounded to 2 decimal places.")
  }

  if (.result != or) {
    fail("Sorry - that's not right. Try using the `epitools` package. Remember
       to round your result to 2 decimal places.")
  }
  pass("Well done!")
})
```

### Interpretation

```{r quiz-crude-or}
quiz(
  caption = NULL,
  question(
    "How would you interpret this crude odds ratio?
    (Select all correct answers)",
    type = "learnr_checkbox",
    answer("Older people are more likely to suffer with SAEs", correct = TRUE),
    answer("Younger people are more likely to suffer with SAEs"),
    answer("The odds of experiencing a SAE is similar in younger and
           older people"),
    answer("The crude odds ratio in younger participants would be 1.0",
      correct = TRUE
    ),
    allow_retry = TRUE
  )
)
```
### Associations with exposre and outcome 

Is diabetes associated with age-group and SAE occurrence? 

```{r diabetes-age-group, echo = TRUE}
epitools::oddsratio.wald(
  x = (df$age_group == "over_45"),
  y = (df$diabetes == "present")
)$measure
```
**Diabetes does seem to be associated with older age.** The odds ratio is 
approximately 2.5 with 95% CI which does not include 1.

Is there an association between diabetes and the occurrence of SAEs?

```{r diabetes-sae, echo = TRUE}
epitools::oddsratio.wald(
  x = (df$diabetes == "present"),
  y = df$sae
)$measure
```
The association here is strong! 

Let's start by checking the stratum specific odds ratios. 

## Stratification

### Stratum-specific ORs

We would like to check whether diabetes might be confounding
the relationship between age group and the occurrence of SAEs. Below is some
more complex code to calculate the same odds ratio in diabetics and 
non-diabetics. See if you can understand what is happening in the code, and 
feel free to ask the course supervisors if you're not sure.

```{r stratum-specific-or, echo = TRUE, warning = FALSE}
df |>
  group_by(diabetes) |>
  summarise(
    n = n(),
    or = epitools::oddsratio.wald(
      x = (.data$age_group == "over_45"),
      y = .data$sae
    )$measure[2]
  )
```

```{r quiz-stratum-specific}
quiz(
  caption = NULL,
  question(
    "Is the effect of age-group on SAE occurence stronger in
              people with diabetes or people without diabetes?",
    answer("People without diabetes"),
    answer("People with diabetes"),
    answer("The effect is very similar in both groups", correct = TRUE),
    answer("Not sure, we need to see the 95% CIs",
      message = "This might be true, but since we're just
                     comparing the two ORs, it's ok to not worry too much about
                     the precision in this instance."
    )
  )
)
```
### Weighted odds ratio estimate

We can use the **Mantel-Haenszel** method to estimate the odds ratio 
for the occurence of SAEs by age-group. 

There are many ways of calculating the MH odds ratio, but in this course, 
we will use the `epiDisplay` package. Since R does not know which level in
our variables we're interested in, we specify them in the function call -
e.g.  `df$age_group == "over_45"`. This is a useful skill to understand in R, 
so please ask the course supervisors if you're not sure what this is doing. 

```{r mh-odds, echo = TRUE}
epiDisplay::mhor(
  sae = df$sae == TRUE,
  age = df$age_group == "over_45",
  diabetes = df$diabetes == "present",
  graph = FALSE
)
```

#### Interpretation

```{r quiz-mh-odds }
quiz(
  caption = NULL,
  question(
    "From the code output, identify the MH Odds Ratio",
    answer(3.36, correct = TRUE),
    answer(3.42),
    answer(13.15),
    answer(12.2),
    random_answer_order = TRUE,
    incorrect = "The MH Odds Ratio is 3.36"
  ),
  question(
    "From the code output, find the upper bound of the
              95% CI",
    answer(6.78, correct = TRUE),
    answer(8.37),
    answer(3.24),
    answer(13.15),
    random_answer_order = TRUE,
    incorrect = "The upper bound of the 95% CI is 6.78"
  ),
  question(
    "From the code output, find the lower bound of the
              95% CI",
    answer(1.66, correct = TRUE),
    answer(1.15),
    answer(0.934),
    answer(0.028547),
    random_answer_order = TRUE,
    incorrect = "The lower bond of the 95% CI is 1.66"
  ),
  question(
    "Compare the crude odds ratio with the MH odds ratio.
              What do you noticce?",
    answer("The MH odds ratio is the same as the crude OR"),
    answer("Diabetes is acting as a confounder", correct = TRUE),
    answer("The crude odds ratio confidence interval is much wider"),
    random_answer_order = TRUE,
    allow_retry = TRUE
  )
)
```

## Logistic regression 

We can use logistic regression to estimate the odds ratio for SAE occurance by
age-group after adjusting for diabetes status. We will use the `broom` package
to make the output more readable.

In the code below, we have provided a ***formula*** to the `glm`
function, along with some information about what type of regression method 
we want to use (`family = binomial(link = "logit")`). While you don't need to 
understand too much about the different types of regression avaialble to us, 
you should realise that the `glm` function can be used with a wide array of
different regression methods. 

Finally, the `tidy` function, exponentiates the returned results and adds 
the 95% confidence interval (`conf.int = TRUE`).

The formula object in R is important to understand when using regression 
methods. Here is the formula in this example:

> sae ~ (age_group == "over_45") + (diabetes == "present"),

There is left-hand side and a right-hand side separated by a *tilde* (`~`).
The left-hand side represents the outcome (SAE occurence in our example), 
and the right-hand side represents the explanatory variables. We will learn 
more about logisic regression in other tutorials and lectures.

You can learn more about the formula in R through this 
[link](https://www.datacamp.com/tutorial/r-formula-tutorial) to a free online 
resource in English. 

```{r lr-odds-ratio, echo = TRUE }
lr_mod <- glm(
  sae ~ (age_group == "over_45") + (diabetes == "present"),
  data = df,
  family = binomial(link = "logit")
) |>
  broom::tidy(
    exponentiate = TRUE,
    conf.int = TRUE
  )
lr_mod
```

#### Interpretation


```{r quiz-lr-odds }
or_est <- round(
  lr_mod$estimate[lr_mod$term == 'age_group == "over_45"TRUE'],
  2
)
quiz(
  caption = NULL,
  question(
    "From the output, what is the odds ratio of SAE occurence by age-group,
    after adjusting for diabetes status?",
    answer(or_est, correct = TRUE),
    answer(9.78),
    answer(14.1),
    answer(19.03),
    random_answer_order = TRUE,
    allow_retry = TRUE
  ),
  question(
    "Does the p-value for this adjusted odds ratio suggest that
              there's strong, moderate or weak evidence for an association
              between age-group and occurance of SAEs?",
    answer("Strong", correct = TRUE),
    answer("Weak"),
    answer("No idea!!"),
    correct = "Correct! The p-value is far smaller than 0.001, suggesting that
    there's stron evidence for an association."
  )
)
```

## Conclusion

In this tutorial, you have calculated a crude and adjusted odds ratio, and 
recognised that diabetes is confounding the effect of age-group on SAE
occurance. We have tried using the Mantel-Haenszel method and have taken 
our first look at logistic regression. As we continue through the course, 
we will use logistic regression to adjust for confounding factors. 











