---
title: "Package development workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Package development workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```
This article provides information that is relevant to anyone contributing
to this package's development. If you notice any errors or have suggestions
to improve the documentation or the package, please get in touch with the 
package creator. 

## Unit testing

To run all package unit tests, run the following code:

```{r}
testthat::auto_test_package()
```

A more robust way of continuously running unit tests is to use the terminal
with `watchexec`. 

First download `watchexec`:

```sh 
brew install watchexec
```

Then run a watch command to re-run unit tests when any R file is updated:

```sh 
watchexec -e R 'Rscript -e "devtools::test()"'
```

## Vignettes

Vignettes are long-form articles which support users through a full workflow 
rather than providing them with detailed documentation about each individual 
package function. 

Vignettes are created in Rmd files and are saved in the `./vignettes` directory.
They need to be manually rendered to ensure they are up-to-date. This is
automatically included in the `devtools::check()` function call if the default
arguments are used. 

During development, it is convenient to review changes to a vignette in the
rendered HTML. This can be set up with `watchexec` using the command below:

```sh 
watchexec -w ./vignettes --exts Rmd "Rscript -e "devtools::install(build_vignettes = TRUE)\""
```

## Tutorials 

This package includes `learnr` interactive tutorials for users to follow. A
seperate article is included in this documentation to show users how to access
tutorials. 

To run tutorials in developent, run the following R command:

```{r} 
devtools::install()
tbmstr::run_tutorial("example")
```
## Internal Rmd 

```r 
watchexec -e Rmd "Rscript -e 'devtools::load_all(); tbmstr::run_quality_check(
    data_path = \"data/test-data/epiinfo/\",
    output_file = file.path(
         system.file(package = \"tbmstr\"),
                                \"dev\", \"quality.html\"
))'"
```
