---
title: "Effects of missing data and misclassification"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
knitr::opts_chunk$set(echo = FALSE)
```


## Effects of missing data 

Since scientific research is dependent on study data, missing values are 
detrimental to the research process. During study implementation, minimising
the amount of missing data is crucial to conducting a rigorous study that 
allows robust conclusions to be drawn.

In this tutorial, you will use R to calculate a mean, an odds ratio, standard
erros and 95% confidence intervals of sample data. At each step, you will see 
how missing data can influence these summary values. 

### Calculating a mean with R 

Write the R code required to calculate the mean of 4, 7, and 1:

```{r manually-calculate-mean, exercise=TRUE}

```
```{r manually-calculate-mean-check}
grade_this({
  if (.result == mean(4, 7, 1)) {
    pass("Well done!")
  }
  fail()
})
```

### Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {

}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaulated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question(
    "Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question(
    "Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```
