---
title: "Effects of missing data and misclassification"
output: 
  learnr::tutorial:
    progressive: true
    df_print: paged
runtime: shiny_prerendered
---

```{r setup, include = FALSE}
library(learnr)
library(gradethis)
library(dplyr)
knitr::opts_chunk$set(
  echo = FALSE,
  exercise.completion = TRUE
)
# define exercise variables
numbers <- c(4, 7, 1)
numbers_missing <- c(4, NA, 1)
```

## Calculating a mean and starndard deviation 

Since scientific research is dependent on study data, missing values are 
detrimental to the research process. During study implementation, minimising
the amount of missing data is crucial to conducting a rigorous study that 
allows robust conclusions to be drawn.

In this tutorial, you will use R to calculate a mean, an odds ratio, standard
erros and 95% confidence intervals of sample data. At each step, you will see 
how missing data can influence these summary values. 

### Calculating a mean with R 

Write the R code required to calculate the mean of `r numbers[1]`, 
`r numbers[2]`, and `r numbers[3]`:

```{r manually-calculate-mean, exercise=TRUE, exercise.lines = 3}

```
```{r manually-calculate-mean-check}
grade_this({
  if (.result == mean(numbers)) {
    pass("Well done! Did you use the `mean` function? Type `?mean` to
         learn more about this important function.")
  }
  fail("Sorry - that's not right. Try using the `mean` function.")
})
```

### Missing values and `mean()`

Notice how the `mean` function does not automatically handle missing values.
Run the code below to see the answer generated, then try to adjust it to
get the correct answer. *Remember that you can read the function documentation
by typing `?mean` into your RStudio terminal*. 

```{r missing-values-mean, exercise = TRUE, exercise.lines = 3}
mean(c(4, NA, 1))
```
```{r missing-values-mean-check}
grade_this({
  if (.result == mean(numbers_missing, na.rm = TRUE)) {
    pass("Well done! Did you read the documentation?")
  }
  fail("Sorry - that's not right. Try reading the help documentation.")
})
```

### Calculating the standard deviation with R

Calculating the standard deviation is more difficult to do manually. Use the
`sd` function to calculate the standard deviation of 4, 7, and 1:

```{r manually-calculate-sd, exercise = TRUE, exercise.lines = 3}

```
```{r manually-calculate-sd-hint}
sd(numbers)
```
```{r manually-calculate-sd-check}
grade_this({
  if (.result == sd(numbers)) {
    pass("Well done! Type `?sd` to learn more.")
  }
  fail("Sorry - that's not right. Try looking at the hint.")
})
```


## Calculating an Odds Ratio

We can use the odds and an odds ratio to compare an outcome with two levels between groups.

### Example data

This example data shows the success and failure of TB treatment in 200 men and women (only the first 10 entries are shown).

```{r odds-setup}
outcomes <- list(
  success_m = 55,
  success_f = 65,
  total = 100
)

male <- data.frame(
  sex = "Male",
  outcome = c(
    rep("Success", outcomes$success_m),
    rep("Failure", outcomes$total - outcomes$success_m)
  )
)

female <- data.frame(
  sex = "Female",
  outcome = c(
    rep("Success", outcomes$success_f),
    rep("Failure", outcomes$total - outcomes$success_f)
  )
)

df <- rbind(male, female)
df <- df[sample(1:nrow(df)), ]

summary <- epitools::oddsratio.wald(
  # include the outcomes
  x = df$outcome,
  # include the exposure which in our case is the participant sex
  y = df$sex
)

or <- summary$measure[2, 1]
lower <- summary$measure[2, 2]
upper <- summary$measure[2, 3]
```

```{r example-data}
knitr::kable(head(df, 10), row.names = FALSE)
```

### Counting success

We can calculate the number of men who were successfully treated with the following code:

```{r success-men, exercise = TRUE, exercise.setup = "odds-setup", exercise.lines = 6 }
df |> 
  filter(sex == "Male", 
         outcome == "Success") |> 
  # nrow counts the number of rows in the data frame after filtering
  nrow()
```

Remember that there were 100 men who started treatment. 

### Counting failure

Can you calculate the number of men who failed treatment? Use the code directly above to help you.

```{r fail-men, exercise = TRUE, exercise.eval = TRUE, exercise.setup = "odds-setup", exercise.lines = 6}

```
```{r fail-men-check}
grade_this({
  if (.result == sum(df$outcome[df$sex == "Male"] == "Failure")) {
    pass("Well done! ")
  }
  fail("Sorry - that's not right. Remember the total number of men receiving treatment is 100.")
})
```


### Calculating odds of success in men

Can you calculate the odds of treatment success in men?

```{r odds-men, exercise = TRUE, exercise.setup = "odds-setup", exercise.lines = 15, exercise.eval = TRUE}

```

```{r odds-men-solution}
success <- df |> 
  filter(sex == "Male", 
         outcome == "Success") |> 
  nrow()

failure <- df |> 
  filter(sex == "Male", 
         outcome == "Failure") |> 
  nrow()

success / failure
```

```{r odds-men-check}
grade_this({
  if (.result == outcomes$success_m / (outcomes$total - outcomes$success_m)) {
    pass("Well done!")
  }
  fail("Sorry - that's not right. Try looking at the solution.")
})
```

### Calculating odds of success in women

Can you calculate the odds of success for women? 

```{r odds-female, exercise = TRUE, exercise.setup = "odds-setup", exercise.lines = 15, exercise.eval = TRUE}
```

```{r odds-female-check}
grade_this({
  if (.result == outcomes$success_f / (outcomes$total - outcomes$success_f)) {
    pass("Well done!")
  }
  fail("Sorry - that's not right. Try looking at the solution.")
})
```
```{r odds-female-solution}
success <- df |> 
  filter(sex == "Female", 
         outcome == "Success") |> 
  nrow()

failure <- df |> 
  filter(sex == "Female", 
         outcome == "Failure") |> 
  nrow()

success / failure
```

### Interpretation?

```{r quiz-odds-success}
quiz(
  question(
    "Which sex is more likely to be treated successfully?",
    answer("Male"),
    answer("Female", correct = TRUE),
    answer("We don't know from the information we have")
  )
)

```


### Odds Ratio

We can quantify the difference in odds of treatment success between men and women using the odds ratio (OR).
From the odds of success in men and women, how can you calculate the odds ratio (OR)? Use the `round` function
to round your answer to 2 decimal places:

```{r odds-ratio, exercise = TRUE, exercise.setup = "odds-setup", exercise.lines = 20, exercise.eval = TRUE}

```

```{r odds-ratio-check}
grade_this({
  if (.result %in% c(round(or, digits = 2),
                     round(1 / or, digits = 2))) {
    pass("Well done!")
  }
  fail("Sorry - that's not right. Try looking at the solution.")
})
```
```{r odds-ratio-solution}
m_success <- df |> 
  filter(sex == "Male", 
         outcome == "Success") |> 
  nrow()

m_failure <- df |> 
  filter(sex == "Male", 
         outcome == "Failure") |> 
  nrow()

m_odds <- m_success / m_failure


f_success <- df |> 
  filter(sex == "Female", 
         outcome == "Success") |> 
  nrow()

f_failure <- df |> 
  filter(sex == "Female", 
         outcome == "Failure") |> 
  nrow()

f_odds <- f_success / f_failure

odds_ratio <- m_odds / f_odds

round(odds_ratio, digits = 2)
```

### Interpretation

```{r quiz-odds-ratio}
quiz(
  question(
    "Which interpretation of the odds ratio is correct?",
    answer("The OR suggests that **men** are treated successfully more
           often than women"),
    answer("The OR suggests that **women** are treated successfully more
           often than men",
           message = "Not the right answer, but you're close."),
    answer("We can't be certain yet because we don't know how precise this result is.",
           correct = TRUE),
    allow_retry = TRUE
  )
)
```

### Precision

We must calculate a 95% Confidence Interval for the Odds Ratio we've calculated. 
To do this manually is too hard, so we can use the `epitools` package:

```{r odds-ratio-95ci, exercise = TRUE, exercise.setup = "odds-setup"}
epitools::oddsratio.wald(
  # include the outcomes
  x = df$outcome,
  # include the exposure which in our case is the participant sex
  y = df$sex
)
```

### Interpretation

```{r quiz-odds-ratio-95ci}
quiz(
  question(
    "From the output above, can you find the lower boundary of the 95% confidence interval?",
    answer("0.6581197"),
    answer("0.1937892"),
    answer("0.372456", correct = TRUE),
    answer("1.0000000"),
    allow_retry = TRUE
  ),
  question(
    "Can you find the upper boundary of the 95% confidence interval?",
    answer("1.162879", correct = TRUE),
    answer("0.1529399"),
    answer("1.648392"),
    answer("3.2038484"),
    allow_retry = TRUE
  ),
  question(
    "Do you expect the p-value to be less than 0.05",
    answer("Yes"),
    answer("No", correct = TRUE),
    allow_retry = TRUE
  ),
  question(
    "Can you find the correct p-value?",
    answer("0.1529399", 
           message = "This is hard - we're going to use the p-value derived from the chi-squared test"),
    answer("0.1937892",
           message = "This is hard - we're going to use the p-value derived from the chi-squared test"),
    answer("0.1489147", correct = TRUE)
  )
)

```

Our final OR is:

> `r glue::glue("**{round(or, digits = 2)} (95%CI {round(lower, digits = 2)}-{round(upper, digits = 2)})**")`

## Effect of misclassification

### Example data

From our previous example data, imaging 10 of the women and 5 of the men 
were given the wrong treatment outcome. This misclassification is mixed amongst
those that actually were successfully treated and those that were not:

```{r misclassification-setup}
outcomes <- list(
  success_m = 53,
  success_f = 71,
  total = 100
)

male <- data.frame(
  sex = "Male",
  outcome = c(
    rep("Success", outcomes$success_m),
    rep("Failure", outcomes$total - outcomes$success_m)
  )
)

female <- data.frame(
  sex = "Female",
  outcome = c(
    rep("Success", outcomes$success_f),
    rep("Failure", outcomes$total - outcomes$success_f)
  )
)

df <- rbind(male, female)
df <- df[sample(1:nrow(df)), ]

```

```{r misclassification-example-data, exercise = TRUE, exercise.setup = "misclassification-setup"}
epitools::tab2by2.test(df$outcome,
             y = df$sex)$x
```



