---
title: "Analysis development"
author: "Jay Achar"
date: "`r Sys.time()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
  officedown::rdocx_document:
    tables:
      layout: autofit
params:
  tables: "Error" 
  plots: "Error" 
  git: "Error"
  text: "Error"
--- 

```{css css, echo=FALSE} 

summary { font-size: 2rem; font-weight: bold; }

.notes {
  color: #00000030;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(flextable)
library(officedown)
library(officer)

# set gtsummary theme options
list("style_number-arg:big.mark" = "") %>%
  set_gtsummary_theme()

if (knitr::is_html_output()) {
  knitr::knit_hooks$set(
    wrap = function(before, options, envir) {
      if (before) {
        paste0("<details><summary>", options$wrap, "</summary>")
      } else {
        paste0("</details>")
      }
    }
  )
}

create_sum_text <- function(
    tab, var, levels, by = NULL, digits = 2,
    pattern = "{total} ({perc}%)") {
  meta <- tab[["meta_data"]]
  df <- meta$df_stats[which(meta$variable == var)][[1]]
  if (!is.null(by)) {
    df <- df[which(df$by %in% by), ]
  }
  target_rows <- which(as.character(df$variable_levels) %in% levels)
  total <- sum(df$n[target_rows])
  denom <- unique(df$N)
  perc <- format(total / denom * 100, digits = digits)
  glue::glue(pattern)
}
```

# Manuscript

The tables, figures and text in this section are most likely to be relevant 
for the manuscript. The manuscript text and notes are included under each 
section so simplify discussion and adjustment. 

## Methods

Baseline clinical and demographic characteristics, and treatment and study
outcomes were described for the treatment cohort. Treatment and study outcomes
were also described by baseline HIV status. The crude effects of baseline CD4
count and HIV treatment regimen were described in participants living with HIV.

Time to culture conversion was described for participants in whom baseline
sputum MGIT culture was positive. The cumulative incidence of culture conversion
was reported by 30 day increments until 120 days.

We described univariable associations between baseline characteristics and
unfavourable study outcomes as crude hazard ratios, 95% confidence intervals and
p-values. Unsuccessful treatment outcomes and death or recurrence during
follow-up were defined as unsuccessful study outcomes. Participants who were
lost to post-treatment follow-up where censored. We then fitted a multivariable
Cox proportional hazard model with a random intercept to account for clustering
by study country. Characteristics with a p<0.1 were evaluated for inclusion.
Eligible characteristics were included in the multivariable model using a
forward stepwise approach. Since very few missing values were present in the 
data set, we used a complete case approach to handling missingness. The effects 
of imputing missing values on model estimates were explored through sensitivity 
analyses.

All statistical analyses were performed with R software (version 4.3.0).

\newpage

<!---BLOCK_LANDSCAPE_START--->

## Treatment cohort description

```{r wrap="Tables & figures", echo = FALSE}
create_text <- function(tab, pattern = "{n} ({p}%)") {
  function(var, level) {
    gtsummary::inline_text(
      tab,
      variable = dplyr::any_of(var),
      level = dplyr::any_of(level),
      pattern = pattern
    )
  }
}

text <- create_text(params$tables$tx_description)
inline <- create_text(params$tables$tx_description, "({n}, {p}%)")

htmltools::h3(
  "Baseline demographic and clinical characteristics of participants"
)
htmltools::hr()
autofit(params$tables$tx_description |>
  gtsummary::as_flex_table())

htmltools::hr()

htmltools::h3(
  "Baseline demographic and clinical characteristics of participants by regimen"
)

autofit(params$tables$tx_description_by_regimen |>
  gtsummary::as_flex_table())
```

<!---BLOCK_LANDSCAPE_STOP--->

<div class="notes">

### Notes

This section is complete. When copying the table into another document, ensure
to address the following style issues: 

* adjust the age-categories into a logical order, 
* the BMI label needs to have a superscript `2`,
* need to add an explanation of abbreviations in the footnote,

**Need to describe participant follow-up time**

</div>

### Manuscript text

After exclusion of (NUMBER) screening failures (Figure 1), 
`r params$tables$tx_description$N` participants from 13 countries were included
in the treatment cohort. More than half of the participants were from Belarus 
`r inline("cntry", "Belarus")` or Ukraine  `r inline("cntry", "Ukraine")`.

Baseline demographic and clinical characteristics are described in Table XX.
The median age was  `r gtsummary::inline_text(params$tables$tx_description, variable = "age", pattern = "{median} (IQR {p25}-{p75})")` with only `r text("age_grp", "<15")` 
participants under 15 years old and `r create_sum_text(params$tables$tx_description, "age_grp", c("<15", "15-24"))` under 24 years old. The large 
majority of participants were male `r text("sex", "Male")`. `r params$text$pregnancy`
Small proportions of participants  reported a history of injecting drug use `r text("idu", "Yes")`, 
homelessness `r text("homeless", "Yes")`, and 
excess alcohol use `r text("alcohol", "Yes")`. Approximately half reported a history of smoking 
`r inline("smok", "Yes")`. Evidence of HIV `r inline("hiv", "Yes")`, HCV 
 `r inline("hcvab", "Seropositive")` and SARS-CoV2 `r inline("covid", "Yes")` 
infection were present in a minority of participants. Baseline positive smear
microscopy `r inline("smear", "Positive")` and cavitatory changes on x-ray 
 `r create_sum_text(params$tables$tx_description, "cav", c("Unilateral",
"Bilateral"), pattern = "({total}, {perc}%)")` were 
detected in over half of participants, while a majority (2136, 81%) had a 
positive baseline MGIT culture result. 

\newpage

## Treatment cohort outcomes

```{r wrap="Tables & figures", echo = FALSE}
eot_text <- create_text(params$tables$tx_outcomes$tbls[[1]])
eot_inline <- create_text(params$tables$tx_outcomes$tbls[[1]], "({n}, {p}%)")
fail_text <- create_text(params$tables$failure_reasons)
fail_inline <- create_text(params$tables$failure_reasons, "({n}, {p}%)")

htmltools::h3(
  "End of treatment outcomes in treatment cohort"
)

params$tables$tx_outcomes |>
  gtsummary::as_flex_table()

htmltools::hr()

htmltools::h3(
  "End of treatment outcomes in treatment cohort by regimen"
)

params$tables$tx_outcomes_by_regimen |>
  gtsummary::as_flex_table()

htmltools::hr()

htmltools::h3(
  "Cumulative probability of TB-free status at specified timepoints",
  "after treatment initiation"
)

params$tables$fail_survival

htmltools::hr()

htmltools::h3(
  "Cumulative probability of TB-free status at specified timepoints",
  "after treatment initiation by regimen"
)

params$tables$fail_survival_by_regimen

htmltools::hr()
params$plots$p1

htmltools::hr()
params$plots$fail_by_regimen

htmltools::hr()
params$plots$p2

htmltools::hr()

htmltools::h3(
  "Cumulative probability of remaining in follow-up after successful treatment
outcome"
)

params$plots$fu_attrition

htmltools::hr()
htmltools::h3(
  "Reasons for treatment failure"
)

params$tables$failure_reasons |>
  gtsummary::as_flex_table()

htmltools::hr()



htmltools::hr()


params$tables$who_fu_outcomes |>
  gtsummary::as_flex_table()
```

<div class="notes">

### Notes 

This section is can be improved. The end of treatment descriptive table is final. 
However, the description of follow-up requires some review. The outstanding
questions are below:

*  Which time points to choose? At this stage I've chosen 301 days to give an 
   accurate description of the end of treatment outcomes. Around 45 participants
   who extended their treatment were given failure outcomes well after 9 months 
   of treatment. They have had their follow-up time manually censored at 300
   days - resulting in the step that's visible in the time to failure graph. 

* Notice how the cumulative probability of TB-free status remains much higher 
  than the `naive` estimate provided in the final table in this section. 

</div>

### Manuscript text

```{r, echo = FALSE}
non_missing <- unique(tables$failure_reasons$meta_data$df_stats[[1]]$N)
reason_count_text <- glue::glue(
  "{non_missing} ",
  "({format(non_missing/unique(tables$failure_reasons$meta_data$df_stats[[1]]$N_obs) * 100, digits = 2)}%)"
)
```
At the end of treatment, `r create_sum_text(params$tables$tx_outcomes$tbls[[1]], "outcome",
c("Cured", "Completed"))` participants were successfully treated and `r eot_text("outcome", "Died")`
died. Out of  `r eot_text("outcome", "Failed")` failed treatment 
episodes, reasons for failure were recorded for `r reason_count_text`. The most
frequently reported reasons were adverse events requiring a treatment change 
`r fail_inline("failure_reasons", "Adverse drug reaction resulting in treatment change")` 
and clinician-driven treatment extensions 
`r fail_inline("failure_reasons", "Treatment extended by clinician")`
and not receiving sufficient treatment during the treatment duration 
`r fail_inline("failure_reasons", "Insufficient treatment received in accepted time")`
(Supplementary table xx). The proportions of participants who were successfully
treated with Regimen 1 (Cs-based regimen) and Regimen 2 (Dlm-based regmein)
were similar (`r create_sum_text(params$tables$tx_outcomes_by_regimen$tbls[[2]], 
"tx_outcome", "Treatment success", "Lfx + Bdq + Lzd + Cfz + Cs")` and `r 
create_sum_text(params$tables$tx_outcomes_by_regimen$tbls[[2]], 
"tx_outcome", "Treatment success", "Lfx + Bdq + Lzd + Cfz + Dlm")`). Treatment
success was more frequent in participants receiving Regimen 3 (without BDq) (`r 
create_sum_text(params$tables$tx_outcomes_by_regimen$tbls[[2]], 
"tx_outcome", "Treatment success", "Lfx + Dlm + Lzd + Cfz")`).

During the post-treatment follow-up period, `r params$text$follow_up$recurrence` 
recurrences and `r params$text$follow_up$death` additional deaths were recorded. 
The cumulative probability of remaining TB-free 9-months, 
15 months and 21 months after treatment initiation was 82%, 80% and 79%
respectively (Table xx and Figure xx).

```{r draft-fup-text-fun, echo = FALSE}
dft <- function(timepoint) {
  df <- params$text$fu_ltfu$stats[[paste0("d", timepoint)]]
  glue::glue("{df$pct}% (95%CI {df$ci$l}-{df$ci$u})")
}
```
From `r params$text$fu_ltfu$total_at_risk` participants who successfully completed treatment, 
 `r params$text$fu_ltfu$stats$d0$events` did not complete any post-treatment 
follow-up. After 6, 11 and 12 months of follow-up, the probability of not having
a negative study outcome and remaining in follow-up was `r dft(180)`, `r dft(330)` and `r dft(365)`
respectively (Appendix Figure). 

\newpage

## Outcomes by HIV status

```{r wrap="Tables & figures", echo = FALSE}
htmltools::h3(
  "End of treament outcomes by baseline HIV status"
)

params$tables$hiv$outcome_by_status |>
  gtsummary::as_flex_table()

htmltools::hr()

htmltools::h3(
  "Cumulative survival probability",
  "after treatment initiation by baseline HIV status"
)

params$plots$p3

htmltools::hr()

htmltools::h3(
  "Description of baseline CD4 count and HIV treatment regimens in ",
  "participants living with HIV"
)
htmltools::hr()
params$tables$hiv$hiv_outcomes |>
  gtsummary::as_flex_table()

htmltools::hr()
params$plots$p4

htmltools::hr()
params$plots$p5
```
<div class="notes">

### Notes

Possible improvements:

* Add risk ratios to compare treatment outcomes by HIV status

</div>

### Manuscript text 

```{r, echo = FALSE}
df <- params$tables$hiv$outcome_by_status$tbls[[1]]$df_by
hiv_cohort <- df[which(df$by == "Yes"), c("n", "p")]
hiv_participants_text <- glue::glue("{hiv_cohort$n} ({round(hiv_cohort$p * 100, 1)}%)")
success <- lapply(
  X = c("Yes", "No"),
  FUN = \(x) {
    create_sum_text(
      tab = params$tables$hiv$outcome_by_status$tbls[[1]],
      var = "outcome",
      by = x,
      levels = c("Cured", "Completed")
    )
  }
) |> setNames(c("Yes", "No"))

death <- lapply(
  X = c("Yes", "No"),
  FUN = \(x) {
    create_sum_text(
      tab = params$tables$hiv$outcome_by_status$tbls[[1]],
      var = "outcome",
      by = x,
      levels = "Died"
    )
  }
) |> setNames(c("Yes", "No"))

ltfu <- lapply(
  X = c("Yes", "No"),
  FUN = \(x) {
    create_sum_text(
      tab = params$tables$hiv$outcome_by_status$tbls[[1]],
      var = "outcome",
      by = x,
      levels = "Lost to follow-up"
    )
  }
) |> setNames(c("Yes", "No"))
```
End-of-treatment responses were less favourable amongst `r hiv_participants_text` 
participants living with HIV, in whom success was recorded for `r success$Yes` 
compared to `r success$No` people without HIV. Death (`r death$Yes` vs 
`r death$No`) and LTFU (`r ltfu$Yes` vs `r ltfu$No`) were also more frequent 
in participants living with HIV compared to those without HIV (Supplementary Table 9).

Amongst `r hiv_participants_text` participants living with HIV, baseline CD4 count 
was available for 245 (91%).  Median baseline CD4 count was 172 (IQR 71-350). 
The majority of participants living with HIV were receiving ART at baseline, with 
dolutegravir-based regimens most frequently used (211, 78%). Almost all
participants living with HIV were receiving co-trimoxazole (254, 97%).

<!---BLOCK_LANDSCAPE_START--->

## Predictors of unsuccessful outcome

```{r wrap="Tables & figures", echo = FALSE}
htmltools::h3(
  "Crude and adjusted analysis of independent predictors of unsuccessful ",
  "study outcome"
)
htmltools::hr()
autofit(params$tables$mv$full)
```
<!---BLOCK_LANDSCAPE_STOP--->

<div class="notes">

### Notes

</div>

### Manuscript text 

```{r, echo = FALSE}
mv_text <- function(var, level, pattern = "(aHR {estimate}; 95%CI {conf.low}-{conf.high})") {
  gtsummary::inline_text(params$tables$mv$adjusted,
    variable = dplyr::all_of(var),
    level = dplyr::all_of(level),
    pattern = pattern
  )
}
```

Crude and adjusted associations between baseline clinical and demographic characteristics 
and unsuccessful study outcome are presented in Table XX. In the adjusted
analysis, increasing age, unemployment `r mv_text("empl_3grp", "Unemployed")`, excessive alcohol use
`r mv_text("alcohol", "Yes")`, history of smoking `r mv_text("smok", "Yes")` , HIV 
positive status `r mv_text("hiv", "Yes")`, presence of bilateral pulmonary 
cavities on x-ray `r mv_text("cav", "Bilateral")`, elevated 
baseline ALT or AST `r mv_text("ast_alt_bin", "Yes")`, and baseline anaemia
`r mv_text("hb_bin", "Yes")` were all positively associated with 
unsuccessful study outcome. Higher BMI `r mv_text("bmi_group", "≥18.5")` 
was negatively associated with unsuccessful study outcome. The positive
association between age and unsuccessful study outcome increased in strength 
with increasing age-group. When compared with participants aged 35-44, those
between 45-54, 55-64, and >64 were `r mv_text("age_grp", "45-54", "{estimate}
(95%CI {conf.low}-{conf.high})")`, `r mv_text("age_grp", "55-64", "{estimate}
(95%CI {conf.low}-{conf.high})")`, and `r mv_text("age_grp", ">64", "{estimate}
(95%CI {conf.low}-{conf.high})")` times more likely to have an unfavourable
study outcome.

\newpage

<!---BLOCK_LANDSCAPE_START--->

## Failure & Death by HIV-specific factors

```{r wrap="Tables & figures", echo = FALSE}
htmltools::h3(
  "Crude associations between baseline HIV treatment and CD4 count and time",
  " to unsuccessful outcome and time to death"
)
htmltools::hr()
params$tables$hiv$failure_death |>
  gtsummary::as_flex_table()
```
```{r, echo = FALSE}
create_hiv_text <- function(
    tab,
    pp = "(HR {estimate}; 95%CI {conf.low}-{conf.high})") {
  return(function(var, level,
                  pattern = pp) {
    gtsummary::inline_text(tab,
      variable = dplyr::all_of(var),
      level = dplyr::all_of(level),
      pattern = pattern
    )
  })
}
failure <- create_hiv_text(params$tables$hiv$failure)
death <- create_hiv_text(params$tables$hiv$death)
```

<!---BLOCK_LANDSCAPE_STOP--->

<div class="notes">

### Notes

</div>

### Manuscript text 

Amongst participants living with HIV `r inline("hiv", "Yes")`, there was
moderate evidence of a positive crude association between not receiving
co-trimoxazole and unsuccessful study outcome `r failure("cpt", "No")` 
and death `r death("cpt", "No")` (Table xxx). Increasing baseline CD4 count appeared to reduce the risk of unsuccessful study
outcome and death. There was moderate evidence of a protective effect in those with 
a CD4 count >500 compared with those with CD4 <=100 on both unsuccessful 
study outcome `r failure("cd4_4grp", "500+")` and death 
`r death("cd4_4grp", "500+")`. Receipt of anti-retroviral therapy at baseline 
was crudely associated with lower risk of death `r death("art", "Yes")`.

\newpage


# Report 

The tables, figures and any text in this section are likely to be relevant 
for the project report. 

## Treatment follow-up in PLHIV

```{r wrap="Tables & figures", echo = FALSE}
htmltools::h3(
  "Treatment follow-up in PLHIV"
)
htmltools::hr()
params$tables$hiv$follow_up |>
  gtsummary::as_flex_table()
```

\newpage

## Outcomes after 12 months follow-up

```{r wrap="Tables & figures", echo = FALSE}
htmltools::h3(
  "Treatment and study oucomes in participants with 12 months follow-up"
)
htmltools::hr()
params$tables$full_fu
```
\newpage

## Probability of study success during follow-up

```{r wrap="Tables & figures", echo = FALSE}
htmltools::h3(
  "Overall probability of successful study outcome during treatment follow-up"
)
htmltools::hr()
params$plots$fu_survival
params$tables$fu_survival
```
\newpage

## Description of deaths

```{r wrap="Tables & figures", echo = FALSE}
htmltools::h3(
  "Study team descriptions of cause of death"
)
htmltools::hr()
params$tables$death_description
```

# Culture conversion risk

<div class="notes">

### Notes

Participants who were defined as culture positive at baseline were included. 
The definition of baseline culture status includes any eligible culture 
results prior to or up to 7 days after starting treatment. Any positive 
eligible culture during this time period would permit participant inclusion 
in these analyses. 

Three definitions of eligible culture have been used to generate the three 
analyses included below:

1. Liquid culture - only cultures using the MGIT960 system are valid,
2. Solid culture - only cultures performed on LJ media are valid,
3. Joint - all cultures, regardless of culture media used, are valid

The validity of cultures, based on the definitions described, have been 
used in each analysis to define the baseline culture status and the culture
conversion event itself. 

For this reason, a participant's inclusion might vary between analysis, as 
well as their culture-conversion status at any given time point.

</div>

## Liquid cultures

```{r wrap="Tables & figures", echo = FALSE}
htmltools::h3(
  "Cumulative incidence of culture conversion by monthly time-points"
)
htmltools::hr()
params$tables$cc_risk$liquid |>
  gtsummary::as_flex_table()

params$plots$conversion$liquid
```
```{r liquid-stats, echo = FALSE}
lq_n <- params$plots$conversion$liquid$data$n.risk[[1]]
total_participants <- params$tables$tx_description$N
lq_median_cc <- gtsummary::inline_text(params$tables$cc_risk$liquid,
  variable = ..overall..,
  column = stat_1_1
)
culture_positive <- paste0(
  lq_n, " (",
  round(lq_n / total_participants * 100, 1), "%)"
)
risk_cc <- lapply(1:4,
  FUN = function(col) {
    statistic <- paste0("stat_", col, "_2")
    params$tables$cc_risk$liquid$table_body[[statistic]]
  }
)
```

### Manuscript text 

At treatment initiation, `r culture_positive` participants were sputum culture positive.
The cumulative risk of culture conversion (IQR) was `r risk_cc[[1]]`, 
`r risk_cc[[2]]`, `r risk_cc[[3]]`, `r risk_cc[[4]]` after 30, 60, 90 and 120
days of treatment respectively. The median days (IQR) to sputum culture conversion 
was `r lq_median_cc`.

## Solid cultures 

```{r wrap="Tables & figures", echo = FALSE}
htmltools::h3(
  "Cumulative incidence of culture conversion by monthly time-points"
)
htmltools::hr()
params$tables$cc_risk$solid |>
  gtsummary::as_flex_table()

params$plots$conversion$solid
```
```{r solid-stats, echo = FALSE}
solid_n <- params$plots$conversion$solid$data$n.risk[[1]]
total_participants <- params$tables$tx_description$N
solid_median_cc <- gtsummary::inline_text(params$tables$cc_risk$solid,
  variable = ..overall..,
  column = stat_1_1
)

culture_positive <- paste0(
  solid_n, " (",
  round(solid_n / total_participants * 100, 1), "%)"
)
risk_cc <- lapply(1:4,
  FUN = function(col) {
    statistic <- paste0("stat_", col, "_2")
    params$tables$cc_risk$solid$table_body[[statistic]]
  }
)
```

### Manuscript text

At treatment initiation, `r culture_positive` participants were sputum culture positive.
The cumulative risk of culture conversion (IQR) was `r risk_cc[[1]]`, 
`r risk_cc[[2]]`, `r risk_cc[[3]]`, `r risk_cc[[4]]` after 30, 60, 90 and 120
days of treatment respectively. The median days (IQR) to sputum culture conversion 
was `r solid_median_cc`.

## All cultures 

```{r wrap="Tables & figures", echo = FALSE}
htmltools::h3(
  "Cumulative incidence of culture conversion by monthly time-points"
)
htmltools::hr()
params$tables$cc_risk$joint |>
  gtsummary::as_flex_table()

params$plots$conversion$joint
```
```{r joint-stats, echo = FALSE}
joint_n <- params$plots$conversion$joint$data$n.risk[[1]]
total_participants <- params$tables$tx_description$N
joint_median_cc <- gtsummary::inline_text(params$tables$cc_risk$joint,
  variable = ..overall..,
  column = stat_1_1
)

culture_positive <- paste0(
  joint_n, " (",
  round(joint_n / total_participants * 100, 1), "%)"
)
risk_cc <- lapply(1:4,
  FUN = function(col) {
    statistic <- paste0("stat_", col, "_2")
    params$tables$cc_risk$joint$table_body[[statistic]]
  }
)
```
### Manuscript text

At treatment initiation, `r culture_positive` participants were sputum culture positive.
The cumulative risk of culture conversion (IQR) was `r risk_cc[[1]]`, 
`r risk_cc[[2]]`, `r risk_cc[[3]]`, `r risk_cc[[4]]` after 30, 60, 90 and 120
days of treatment respectively. The median days (IQR) to sputum culture conversion 
was `r joint_median_cc`.

\newpage

# Debug

Git branch: `r params$git$branch`  
Git commit: `r params$git$hash`  
Link: [Github](`r paste0("https://github.com/JayAchar/tbmstr/commit/", params$git$hash)`)

