---
title: "Sensitivity analyses"
author: "Jay Achar"
date: "`r Sys.time()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
params:
  git: "Error"
  data: "Error"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(flextable)
```
# Characteristics by Follow-up status

Below is a table which compares the baseline characteristics of participants
who:

* Successfully completed treatment and,
* Either completed or did not complete 600 days in the study

```{r}
params$data$characteristics_by_fu
```

# Exclude Ukraine

Since the war has affected the care provided in Ukraine, it's worth excluding
their results to see the effect on the overall study results:

## End of Treatment Outcome

```{r}
params$data$exclude_ukr$eot_outcome
```

## End of Study Outcome

```{r}
params$data$exclude_ukr$eos_outcome
```


# Debug

Git branch: `r params$git$branch`  
Git commit: `r params$git$hash`  
Link: [Github](`r paste0("https://github.com/JayAchar/tbmstr/commit/", params$git$hash)`)

