---
title: "The importance of confounding"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
description: 'Understanding and handling confounding'
---

```{r setup, include = FALSE}
library(learnr)
library(gradethis)
library(dplyr)
library(epitools)
knitr::opts_chunk$set(
  echo = FALSE
)
tutorial_options(
  exercise.completion = TRUE
)
# create example data
total <- 500L
sex <- c(rep("Male", total * 0.5),
         rep("Female", total * 0.5))

smear <- c(
  rep("Negative", total * 0.05),
  rep("Positive", total * 0.45),
  rep("Negative", total * 0.35),
  rep("Positive", total * 0.15)
)

simulate <- \() {
  set.seed(5555)
  beta0 <- 1.1
  beta1 <- log(2)
  beta2 <- log(0.3)
  pi_x <- exp(beta0 + beta1 * (sex == "Female") + beta2 * (smear == "Positive")) / 
    (1 + exp(beta0 + beta1 * (sex == "Female") + beta2 * (smear == "Positive")))
  outcome <- rbinom(n = length(sex), size = 1, prob = pi_x)
  outcome <- ifelse(outcome == 1, "Success", "Failure")
  data.frame(sex, smear, outcome)
}

df <- simulate()

df <- df[sample(1:nrow(df)), ]

crude <- epitools::oddsratio.wald(
  x = df$sex,
  y = df$outcome
)
```

## Confounding

Confounding is all about alternative explanations for an effect being seen. 

In this tutorial, we will use TB treatment outcome, sex and baseline smear 
microscopy status variables. The example data will include information from
`r total * 0.5` men and `r total * 0.5` women. The first 10 rows are shown below:

### Example data

```{r example-data}
knitr::kable(head(df, 10), row.names = FALSE)
```

```{r outcome-by-sex}
table(df$sex, df$outcome)
```

### Crude association between sex and treatment outcome

Can you calculate the odds ratio for treatment success in men compared to women?
**Round your result to 2 decimal places**

```{r odds-sex-outcome, exercise = TRUE, exercise.lines = 7}

```
```{r odds-sex-outcome-solution}
result <- epitools::oddsratio.wald(
  x = df$sex,
  y = df$outcome
)

round(result$measure[2], 
      digits = 2)
```

> Crude association with 95% CI: `r glue::glue("**{round(crude$measure[2], digits = 2)} (95%CI {round(crude$measure[4], digits = 2)}-{round(crude$measure[6], digits = 2)})**")`

```{r odds-sex-outcome-check}
or <- round(epitools::oddsratio.wald(
  x = df$sex,
  y = df$outcome
)$measure[2], 2)

grade_this({
  if (.result == or) {
    pass("Well done! ")
  }
  fail("Sorry - that's not right. Try using the `epitools` package. Remember
       to round your result to 2 decimal places.")
})
```

### Effect of baseline smear status

To be considered a confounder, the baseline smear status needs to be associated
with the treatment outcome and sex. 

```{r sex-by-outcome-stratified}
smear <- c("Positive", "Negative")
lapply(
  smear,
  \(level) {
    strata <- df[which(df$smear == level), ]
    table(strata$sex, strata$outcome)
  }
) |> setNames(smear)
```


Can you calculate the odds ratio between treatment outcome and baseline 
smear status? **Round your result to 2 decimal places**

```{r smear-outcome}
table(df$smear, df$outcome)
```


```{r odds-smear-outcome, exercise = TRUE, exercise.lines = 7}

```
```{r odds-smear-outcome-hint}
"Can you copy and adjust the previous example?"
```
```{r odds-smear-outcome-solution}
result <- epitools::oddsratio.wald(
  x = df$smear,
  y = df$outcome
)

round(result$measure[2], 
      digits = 2)
```
```{r odds-smear-outcome-check}
or <- round(epitools::oddsratio.wald(
  x = df$smear,
  y = df$outcome
)$measure[2], 2)

grade_this({
  if (.result == or) {
    pass("Well done! ")
  }
  fail("Sorry - that's not right. Try using the `epitools` package. Remember
       to round your result to 2 decimal places.")
})
```


Can you calculate the odds ratio between sex and baseline smear status?
**Round your result to 2 decimal places**

```{r sex-smear}
table(df$sex, df$smear)
```


```{r odds-smear-sex, exercise = TRUE, exercise.lines = 7}

```
```{r odds-smear-sex-hint}
"Can you copy and adjust the previous example?"
```
```{r odds-smear-sex-solution}
result <- epitools::oddsratio.wald(
  x = df$smear,
  y = df$sex
)

round(result$measure[2], 
      digits = 2)
```
```{r odds-smear-sex-check}
or <- round(epitools::oddsratio.wald(
  x = df$smear,
  y = df$sex
)$measure[2], 2)

grade_this({
  if (.result == or) {
    pass("Well done! ")
  }
  fail("Sorry - that's not right. Try using the `epitools` package. Remember
       to round your result to 2 decimal places.")
})
```

### Interpretation

```{r quiz-confounding}
quiz(
  caption = NULL,
  question(
    "Is there an association between smear status and treatment outcome?",
    answer("Yes", correct = TRUE),
    answer("No"),
    answer("Not sure", message = "Have another look at the table showing baseline smear
           status and treatment outcome. Are the successes and failures equally distributed?"),
    allow_retry = TRUE
  ),
  question(
    "Is there an association between smear status and sex?",
    answer("Yes", correct = TRUE),
    answer("No"),
    answer("Not sure", message = "Have another look at the table showing baseline smear
           status and tsex Are the smear positive and negative results equally distributed
           across sexes?"),
    allow_retry = TRUE
  )
)
```

## Controlling for confounding

There are multiple methods of controlling for confounding in epidemiological
studies. 

```{r quiz-controlling}
wrong <- "No, this is a method that can be used when designing a study."
quiz(
  caption = NULL,
  question(
    "Choose all methods to control for confounding in the analysis (multiple choices)",
    answer("Restriction", message = wrong),
    answer("Stratification", correct = TRUE),
    answer("Randomisation", message = wrong),
    answer("Regression modelling", correct = TRUE),
    answer("Matching", message = wrong)
  )
)
```

Below is the crude association we calculatd in the earlier part of this tutorial:

> Crude odds ratio with 95% CI: `r glue::glue("**{round(crude$measure[2], digits = 2)} (95%CI {round(crude$measure[4], digits = 2)}-{round(crude$measure[6], digits = 2)})**")`

Using multivariable logistic regression, we can adjust for the confounding
effect of smear to give us a more accurate estimate of the association between
sex and treatment outcome: 

```{r regression, echo = TRUE}
result <- glm((outcome == "Success") ~ sex + smear,
    data = df,
    family = binomial(link = "logit")) |> 
  broom::tidy(
    exponentiate = TRUE,
    conf.int = TRUE
  )

or <- result[which(result$term == "sexMale"), ]

result
```

The middle row of the output shows the sex and treatment outcome. The adjusted odds ratio is 
shown below:

> Adjusted odds ratio with 95% CI: `r glue::glue("**{round(or$estimate, digits = 2)} (95%CI {round(or$conf.low, digits = 2)}-{round(or$conf.high, digits = 2)})**")`




















