---
title: "Analysis development"
author: "Jay Achar"
date: "`r Sys.time()`"
output: 
  - html_document
  - officedown::rdocx_document
params:
  tables: "Error" 
  plots: "Error" 
--- 

```{css css, echo=FALSE} 

summary { font-size: 2rem; font-weight: bold; }

h2 {
  font-size: 2rem;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(flextable)
library(officedown)
library(officer)

# set gtsummary theme options
list("style_number-arg:big.mark" = "") %>%
  set_gtsummary_theme()

if (knitr::is_html_output()) {
  knitr::knit_hooks$set(
    wrap = function(before, options, envir) {
      if (before) {
        paste0("<details><summary>", options$wrap, "</summary>")
      } else {
        paste0("</details>")
      }
    }
  )
}

create_sum_text <- function(
    tab, var, levels, by = NULL, digits = 2,
    pattern = "{total} ({perc}%)") {
  meta <- tab[["meta_data"]]
  df <- meta$df_stats[which(meta$variable == var)][[1]]
  if (!is.null(by)) {
    df <- df[which(df$by %in% by), ]
  }
  target_rows <- which(as.character(df$variable_levels) %in% levels)
  total <- sum(df$n[target_rows])
  denom <- unique(df$N)
  perc <- format(total / denom * 100, digits = digits)
  glue::glue(pattern)
}
```
## Methods

Baseline clinical and demographic characteristics, and treatment and study
outcomes were described for the treatment cohort. Treatment and study outcomes
were also described by baseline HIV status. The crude effects of baseline CD4
count and HIV treatment regimen were described in participants living with HIV.

Time to culture conversion was described for participants in whom baseline
sputum MGIT culture was positive. The cumulative incidence of culture conversion
was reported by 30 day increments until 120 days.

We described univariable associations between baseline characteristics and
unfavourable study outcomes as crude hazard ratios, 95% confidence intervals and
p-values. Unsuccessful treatment outcomes and death or recurrence during
follow-up were defined as unsuccessful study outcomes. Participants who were
lost to post-treatment follow-up where censored. We then fitted a multivariable
Cox proportional hazard model with a random intercept to account for clustering
by study country. Characteristics with a p<0.1 were evaluated for inclusion.
Eligible characteristics were included in the multivariable model using a
forward stepwise approach. Since very few missing values were present in the 
data set, we used a complete case approach to handling missingness. The effects 
of imputing missing values on model estimates were explored through sensitivity 
analyses.

All statistical analyses were performed with R software (version 4.3.0).

\newpage

```{r wrap="Treatment cohort description", echo = FALSE}
create_text <- function(tab, pattern = "{n} ({p}%)") {
  function(var, level) {
    gtsummary::inline_text(
      tab,
      variable = dplyr::any_of(var),
      level = dplyr::any_of(level),
      pattern = pattern
    )
  }
}

text <- create_text(params$tables$tx_description)
inline <- create_text(params$tables$tx_description, "({n}, {p}%)")

htmltools::h2(
  "Baseline demographic and clinical characteristics of participants"
)
htmltools::hr()
autofit(params$tables$tx_description |>
  gtsummary::as_flex_table())
```
## Notes 

This section is complete. When copying the table into another document, ensure
to address the following style issues: 

* adjust the age-categories into a logical order, 
* the BMI label needs to have a superscript `2`,
* the country order should be alphabetical,
* need to add an explanation of abbreviations in the footnote,

**Need to add baseline MGIT culture status to table.**

**Need to describe participant follow-up time**

## Manuscript text

After exclusion of (NUMBER) screening failures (Figure 1), 2636
participants from 13 countries were included in the treatment 
cohort. More than half of the participants were from Belarus `r inline("cntry", "Belarus")` or Ukraine
 `r inline("cntry", "Ukraine")`.

Baseline demographic and clinical characteristics are described in Table XX.
The median age was XX years (IQR: ) with only `r text("age_grp", "<15")` 
participants under 15 years old and `r create_sum_text(params$tables$tx_description, "age_grp", c("<15", "15-24"))` under 24 years old. The large 
majority of participants were male `r text("sex", "Male")`. Small proportions of participants 
reported a history of injecting drug use `r text("idu", "Yes")`, 
homelessness `r text("homeless", "Yes")`, and 
excess alcohol use `r text("alcohol", "Yes")`. Approximately half reported a history of smoking 
`r inline("smok", "Yes")`. Evidence of HIV `r inline("hiv", "Yes")`, HCV 
 `r inline("hcvab", "Seropositive")` and SARS-CoV2 `r inline("covid", "Yes")` 
infection were present in a minority of participants. Baseline positive smear
microscopy `r inline("smear", "Positive")` and cavitatory changes on x-ray 
 `r create_sum_text(params$tables$tx_description, "cav", c("One-sided",
"Two-sided"), pattern = "({total}, {perc}%)")` were 
detected in over half of participants, while a majority (2136, 81%) had a 
positive baseline MGIT culture result.

\newpage

```{r wrap="Treatment cohort outcomes", echo = FALSE}
eot_text <- create_text(params$tables$tx_outcomes)
eot_inline <- create_text(params$tables$tx_outcomes, "({n}, {p}%)")

fail_text <- create_text(params$tables$failure_reasons)
fail_inline <- create_text(params$tables$failure_reasons, "({n}, {p}%)")

htmltools::h2(
  "End of treatment outcomes in treatment cohort"
)

params$tables$tx_outcomes |>
  gtsummary::as_flex_table()

htmltools::hr()

htmltools::h2(
  "Cumulative probability of TB-free status at specified timepoints",
  "after treatment initiation"
)

params$tables$fail_survival

htmltools::hr()
params$plots$p1

htmltools::hr()
params$plots$p2

htmltools::hr()
htmltools::h2(
  "Reasons for treatment failure"
)

params$tables$failure_reasons |>
  gtsummary::as_flex_table()

htmltools::hr()



htmltools::hr()


params$tables$who_fu_outcomes |>
  gtsummary::as_flex_table()
```

## Notes 

This section is can be improved. The end of treatment descriptive table is final. 
However, the description of follow-up requires some review. The outstanding
questions are below:

*  Which time points to choose? At this stage I've chosen 301 days to give an 
   accurate description of the end of treatment outcomes. Around 45 participants
   who extended their treatment were given failure outcomes well after 9 months 
   of treatment. They have had their follow-up time manually censored at 300
   days - resulting in the step that's visible in the time to failure graph. 

* Notice how the cumulative probability of TB-free status remains much higher 
  than the `naive` estimate provided in the final table in this section. 

**Remove unused levels from the `Reasons for failure` table**

## Manuscript text

```{r, echo = FALSE}
non_missing <- unique(tables$failure_reasons$meta_data$df_stats[[1]]$N)
reason_count_text <- glue::glue(
  "{non_missing} ",
  "({format(non_missing/unique(tables$failure_reasons$meta_data$df_stats[[1]]$N_obs) * 100, digits = 2)}%)"
)
```
At the end of treatment, `r create_sum_text(params$tables$tx_outcomes, "outcome",
c("Cured", "Completed"))` participants were successfully treated and `r eot_text("outcome", "Died")`
died. Out of  `r eot_text("outcome", "Failed")` failed treatment 
episodes, reasons for failure were recorded for `r reason_count_text`. The most
frequently reported reasons were adverse events requiring a treatment change 
`r fail_inline("failure_reasons", "Adverse drug reaction resulting in treatment change")` 
and clinician-driven treatment extensions 
`r fail_inline("failure_reasons", "Treatment extended by clinician")`
and not receiving sufficient treatment during the treatment duration 
`r fail_inline("failure_reasons", "Insufficient treatment received in accepted time")`
(Supplementary table xx). 

During the post-treatment follow-up period, 25 recurrences and 44 additional 
deaths were recorded. The cumulative probability of remaining TB-free 9-months, 
15 months and 21 months after treatment initiation was 82%, 80% and 79%
respectively (Table xx and Figure xx).

\newpage

```{r wrap="Outcomes by HIV status", echo = FALSE}
htmltools::h2(
  "End of treament outcomes by baseline HIV status"
)

params$tables$hiv$outcome_by_status |>
  gtsummary::as_flex_table()

htmltools::hr()

htmltools::h2(
  "Cumulative survival probability",
  "after treatment initiation by baseline HIV status"
)

params$plots$p3

htmltools::hr()

htmltools::h2(
  "Description of baseline CD4 count and HIV treatment regimens in ",
  "participants living with HIV"
)
htmltools::hr()
params$tables$hiv$hiv_outcomes |>
  gtsummary::as_flex_table()

htmltools::hr()
params$plots$p4

htmltools::hr()
params$plots$p5
```
## Notes

Possible improvements:

* Add risk ratios to compare treatment outcomes by HIV status


## Manuscript text 

```{r, echo = FALSE}
df <- params$tables$hiv$outcome_by_status$df_by
hiv_participants <- df[which(df$by == "Yes"), "n"]
success <- lapply(
  X = c("Yes", "No"),
  FUN = \(x) {
    create_sum_text(
      tab = params$tables$hiv$outcome_by_status,
      var = "outcome",
      by = x,
      levels = c("Cured", "Completed")
    )
  }
) |> setNames(c("Yes", "No"))
```
End of treatment responses were less favourable in the `r hiv_participants` participants 
living with HIV. Treatment success was `r success$Yes` compared to `r success$No` in people without HIV.
Death (xx) and LTFU (xx) were also more frequent (Supplementary table xx, 
Sepplementary figure xx).  

Amongst `r hiv_participants` participants living with HIV, baseline CD4 count 
was available for 245 (91%).  Median baseline CD4 count was 172 (IQR 71-350). 
The majority of participants living with HIV were receiving ART at baseline, with 
dolutegravir-based regimens most frequently used (211, 78%). Almost all
participants living with HIV were receiving cotrimoxazole (254, 97%).

<!---BLOCK_LANDSCAPE_START--->

```{r wrap="Predictors of unsuccessful outcome", echo = FALSE}
htmltools::h2(
  "Crude and adjusted analysis of independent predictors of unsuccessful ",
  "study outcome"
)
htmltools::hr()
autofit(params$tables$mv$full)
```
<!---BLOCK_LANDSCAPE_STOP--->

## Notes


## Manuscript text 

```{r, echo = FALSE}
mv_text <- function(var, level, pattern = "(HR {estimate}; 95%CI {conf.low}-{conf.high})") {
  gtsummary::inline_text(tables$mv$adjusted,
    variable = dplyr::all_of(var),
    level = dplyr::all_of(level),
    pattern = pattern
  )
}
```

Crude and adjusted associations between baseline clinical and demographic characteristics 
and unsuccessful study outcome are presented in Table XX. Following adjustment 
for strongly associated factors and clustering by study country, increasing
age, unemployment `r mv_text("empl_3grp", "Unemployed")`, excessive alcohol use
`r mv_text("alcohol", "Yes")`, history of smoking `r mv_text("sm_fact", "<20 pack-years")` , HIV 
positive status `r mv_text("hiv", "Yes")`, presence of bilateral pulmonary 
cavities on x-ray `r mv_text("cav", "Two-sided")`, elevated 
baseline ALT or AST `r mv_text("ast_alt_bin", "Yes")`, and baseline anaemia
`r mv_text("hb_bin", "Yes")` were all positively associated with 
unsuccessful study outcome. Higher BMI `r mv_text("bmi_group", "≥18.5")` 
was negatively associated with unsuccessful study outcome. The positive
association between age and unsuccessful study outcome increased in strength 
with increasing age-group. When compared with participants aged 35-44, those
between 45-54, 55-64, and >64 were `r mv_text("age_grp", "45-54", "{estimate}
(95%CI {conf.low}-{conf.high})")`, `r mv_text("age_grp", "55-64", "{estimate}
(95%CI {conf.low}-{conf.high})")`, and `r mv_text("age_grp", ">64", "{estimate}
(95%CI {conf.low}-{conf.high})")` times more likely to have an unfavourable
study outcome.

\newpage

```{r wrap="Treatment follow-up in PLHIV", echo = FALSE}
htmltools::h2(
  "Treatment follow-up in PLHIV"
)
htmltools::hr()
params$tables$hiv$follow_up |>
  gtsummary::as_flex_table()
```

<!---BLOCK_LANDSCAPE_START--->

```{r wrap="Failure & Death by HIV-specific factors", echo = FALSE}
htmltools::h2(
  "Crude associations between baseline HIV treatment and CD4 count and time",
  " to unsuccessful outcome and time to death"
)
htmltools::hr()
params$tables$hiv$failure_death |>
  gtsummary::as_flex_table()
```

<!---BLOCK_LANDSCAPE_STOP--->

```{r wrap="Culture conversion risk", echo = FALSE}
htmltools::h2(
  "Cumulative incidence of culture conversion by monthly time-points"
)
htmltools::hr()
params$tables$cc_risk

params$plots$conversion
```

\newpage

```{r wrap="Outcomes after 12 months follow-up", echo = FALSE}
htmltools::h2(
  "Treatment and study oucomes in participants with 12 months follow-up"
)
htmltools::hr()
params$tables$full_fu
```
\newpage

```{r wrap="Probability of study success during follow-up", echo = FALSE}
htmltools::h2(
  "Overall probability of successful study outcome during treatment follow-up"
)
htmltools::hr()
params$plots$fu_survival
params$tables$fu_survival
```
\newpage

```{r wrap="Description of deaths", echo = FALSE}
htmltools::h2(
  "Study team descriptions of cause of death"
)
htmltools::hr()
params$tables$death_description
```

